Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-07-09 manuscript/manuscript.Rmd
updates ms with new references

where $k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})$ is the population kernal that describes all possible transitions from size $u$ to $v$ and $\bar{\bold{w_{j}}}$ is a vector of estimates of average crowding experienced from all other species by a genet of size $u_j$ and species $j$. The integral is evaluated over all possible sizes between predefined lower (_L_) and upper (_U_) size limits that extend beyond the range of observed genet sizes.

The population kernal is defined as the joint contributions of survival (_S_), growth (_G_), and recruitment (_R_):

@@

Equilibrium cover from both models was sensitive to climate (Fig. 3a-d). The IPM projected percent changes in equilibrium cover from `r round(min(subset(meanparam_agg, species=="BOGR" & model=="IPM")["meandiff"])*100)` - `r round(max(subset(meanparam_agg, species=="BOGR" & model=="IPM")["meandiff"])*100)`% for *B. gracilis*, `r round(min(subset(meanparam_agg, species=="HECO" & model=="IPM")["meandiff"])*100)` - `r round(max(subset(meanparam_agg, species=="HECO" & model=="IPM")["meandiff"])*100)`% for *H. comata*, `r round(min(subset(meanparam_agg, species=="PASM" & model=="IPM")["meandiff"])*100)` - `r round(max(subset(meanparam_agg, species=="PASM" & model=="IPM")["meandiff"])*100)`% for *P. smithii*, and `r round(min(subset(meanparam_agg, species=="POSE" & model=="IPM")["meandiff"])*100)` - `r round(max(subset(meanparam_agg, species=="POSE" & model=="IPM")["meandiff"])*100)`% for *P. secunda*. The QBM projected opposite and greater percent changes in equilibrium cover for *B. gracilis* (`r round(min(subset(meanparam_agg, species=="BOGR" & model=="QBM")["meandiff"])*100)` - `r round(max(subset(meanparam_agg, species=="BOGR" & model=="QBM")["meandiff"])*100)`%) and *H. comata* (`r round(min(subset(meanparam_agg, species=="HECO" & model=="QBM")["meandiff"])*100)` - `r round(max(subset(meanparam_agg, species=="HECO" & model=="QBM")["meandiff"])*100)`%; Fig. 3a-b). For *P. smithii*, the QBM projected opposite changes in equilibrium cover than the IPM, but of similar magnitude (`r round(min(subset(meanparam_agg, species=="PASM" & model=="QBM")["meandiff"])*100)` - `r round(max(subset(meanparam_agg, species=="PASM" & model=="QBM")["meandiff"])*100)`%; Fig. 3c). *P. secunda* was the only species that the IPM and QBM made projections of the same sign and somewhat similar magnitude (Fig. 3d).

The response of a population to climate change is a result of the aggregate effects of climate on individual vital rates. Since the IPM approach relies on vital rate regressions, we were able to quantify the sensitivity of each vital rate in isolation and in pairwise combinations. Species showed similar trends (Fig. 3e-h). Growth was the most sensitive vital rate for all species, showing a negative response to increased precipitation, and stronger positive response to increased temperature, and a mostly positive response when both climate factors are increased (Fig. 3e-h). *B. gracilis* survival rates were sensitive to temperature, resulting in an increase in plant cover under increased temperature (Fig. 3e). In isolation, recruitment and survival were insensitive to climate factors for *H. comata* (Fig. 3f). Survival and recruitment of *P. smithii* were both sensitive, negatively, to temperature and precipitation (Fig. 3g). *P. secunda* equilibrium cover was sensitive to the climate effects on survival and recruitment, showing a negative effect on both vital rates for increased precipition, but a strong positive effect on survival with increased temperature (Fig. 3h). The climate impact of recruitment on equilibrium cover was negative for precipitation and temperature increases (Fig. 3h).

@@

These results lead us to conclude that detailed demographic data is necessary to forecast climate change impacts on plant populations. This is unwelcome news since this data is difficult to collect and the models built on such data are of little use to land managers that make decisions at scales beyond that of traditional population models [@Queenborough2011]. While density-structured approaches may fail when climate covariates are considered, there are other alternatives. For example, @Clark2011a use Forest Inventory and Analysis (FIA) data to parameterize a population model with multiple vital rates and climate dependence. Another example are distributed efforts like PlantPopNet (http://plantago.plantpopnet.com) that will allow researchers to estimate variation around climate responses for widespread species. Lastly, we foresee new approaches on the horizon that leverage photo/video of plots and advanced object recognition algorithms to streamline plant mapping and digitizing efforts.

### Forecasting the future, and the future of forecasting
Our goal was not make any explicit forecast for the future state of these populations based on predicted climate change. But our results highlight the state of affairs in ecology when it comes to forecasting the impacts of climate change. The analysis we conducted here could be considered, with some exceptions of course, at the forefront of ecological forecasting in terms of the statistical approach employed (hierarchical Bayesian), the type of population model we used (stochastic IPM with parameter uncertainty), and the amount of data we had at our disposal (14 years of individual-level data). Yet, model predictions proved so uncertain that any forecast, when bounded with model and parameter uncertainty, would be at best not useful and at worst meaningless. How might we improve on this state of affairs?

First, forecasts could be improved by matching the spatial scale of predictor variables with the spatial scale of observations. One of the major limitations of the models we fit here is that the climate data are at a much larger scale than the plot and individual level observations of plant cover and size. Climate covariates only vary by year, with no spatial variability within years. Thus, even if we fit models to individual level data, we are missing the key interaction point between weather and individual plants [@Clark2011b].

Second, accurately detecting climate signals will take even longer time series. Recent theoretical work on detecting climate signals in noisy data suggests that even advanced approaches to parameter fitting like LASSO, functional linear models (splines), and random forest require 20-25 year time series (Teller et al., in review). Alternatively, as we suggest above, Teller et al. (in review) also find that matching the scale of the response and predictors improves estimate precision.

Third, ecologists as a community need to get serious about reporting uncertainty. There is a strong culture around explicitly considering model uncertainty, but parameter uncertainty is often ignored. In some cases this is because the easiest statistical methods do no make propogating parameter uncertainty easy. Even using Bayesian approaches that allow straightforward integration of model fitting and forecasting [@Hobbs2015] is not simple when using modeling approaches like integral projection models that separate the model fitting and simulation stages [@Rees2009]. However, as we have done here, it is still possible to include parameter uncertainty by drawing parameter values from MCMC iterations, taking care to draw all parameters from the same chain and iteration to account for their correlations. Only by being honest about our forecasts can we begin to produce better ones.

Conclusions
-----------
This work is not a critique of density-structured population models. In some cases and for certain species, population models based on aggregated data may prove useful and unbiased. However, our work here is the first comparison, to our knowledge, of population models based on individual and aggregated data in a climate change context. Our results confirm theoretical arguments [@Clark2011b] and empirical evidence [@Clark2011; @Clark2012] that individual responses are critical to predicting species' responses to climate change. Thus, forecasts from aggregate level models should be viewed with caution and should never be unaccompianed by uncertainty. Given the importance of demographic data and its current difficulty to collect, we advocate for research on modern methods to collect demographic data more efficiently across environmental gradients in space and time. For example, recent advances in object recognition algorithms [@Liu2014] provide new opportunities to speed up data collection and digital processing.

@@

Acknowledgments
---------------
This work was funded by the National Science Foundation through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040). We thank the original mappers of the permanent quadrats in Montana and the digitizers in the Adler lab, without whom this work would not have been possible. Informal conversations with Stephen Ellner, Giles Hooker, Robin Snyder, and a series of meetings among the Adler and Weecology labs at USU sharpened our thinking. Compute, storage and other resources from the Division of Research Computing in the Office of Research and Graduate Studies at Utah State University are gratefully acknowledged.

>>>>

where $k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})$ is the population kernel that describes all possible transitions from size $u$ to $v$ and $\bar{\bold{w_{j}}}$ is a vector of estimates of average crowding experienced from all other species by a genet of size $u_j$ and species $j$. The integral is evaluated over all possible sizes between predefined lower (_L_) and upper (_U_) size limits that extend beyond the range of observed genet sizes.

The population kernel is defined as the joint contributions of survival (_S_), growth (_G_), and recruitment (_R_):

@@

Equilibrium cover from both models was sensitive to climate (Fig. 3a-d). The IPM projected percent changes in equilibrium cover from `r round(min(subset(meanparam_agg, species=="BOGR" & model=="IPM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="BOGR" & model=="IPM")["meandiff"])*100)`% for *B. gracilis*, `r round(min(subset(meanparam_agg, species=="HECO" & model=="IPM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="HECO" & model=="IPM")["meandiff"])*100)`% for *H. comata*, `r round(min(subset(meanparam_agg, species=="PASM" & model=="IPM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="PASM" & model=="IPM")["meandiff"])*100)`% for *P. smithii*, and `r round(min(subset(meanparam_agg, species=="POSE" & model=="IPM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="POSE" & model=="IPM")["meandiff"])*100)`% for *P. secunda*. The QBM projected opposite and greater percent changes in equilibrium cover for *B. gracilis* (`r round(min(subset(meanparam_agg, species=="BOGR" & model=="QBM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="BOGR" & model=="QBM")["meandiff"])*100)`%) and *H. comata* (`r round(min(subset(meanparam_agg, species=="HECO" & model=="QBM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="HECO" & model=="QBM")["meandiff"])*100)`%; Fig. 3a-b). For *P. smithii*, the QBM projected opposite changes in equilibrium cover than the IPM, but of similar magnitude (`r round(min(subset(meanparam_agg, species=="PASM" & model=="QBM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="PASM" & model=="QBM")["meandiff"])*100)`%; Fig. 3c). *P. secunda* was the only species that the IPM and QBM made projections of the same sign and somewhat similar magnitude (`r round(min(subset(meanparam_agg, species=="POSE" & model=="QBM")["meandiff"])*100)` to `r round(max(subset(meanparam_agg, species=="POSE" & model=="QBM")["meandiff"])*100)`%; Fig. 3d).

The response of a population to climate change is a result of the aggregate effects of climate on individual vital rates. Since the IPM approach relies on vital rate regressions, we were able to quantify the sensitivity of each vital rate in isolation and in pairwise combinations. Species showed similar trends (Fig. 3e-h). Growth was the most sensitive vital rate for all species, showing a negative response to increased precipitation, and stronger positive response to increased temperature, and a mostly positive response when both climate factors are increased (Fig. 3e-h). *B. gracilis* survival rates were sensitive to temperature, resulting in an increase in plant cover under increased temperature (Fig. 3e). In isolation, recruitment and survival were insensitive to climate factors for *H. comata* (Fig. 3f). Survival and recruitment of *P. smithii* were both sensitive, negatively, to temperature and precipitation (Fig. 3g). *P. secunda* equilibrium cover was sensitive to the climate effects on survival and recruitment, showing a negative effect on both vital rates for increased precipition, but a strong positive effect on survival with increased temperature (Fig. 3h). The climate impact of recruitment on equilibrium cover was negative for precipitation and temperature increases (Fig. 3h). At least two of three vital rates were sensitive to climate for each species (Fig. 3).

@@

These results lead us to conclude that detailed demographic data is necessary to forecast climate change impacts on plant populations. This is unwelcome news since this data is difficult to collect and the models built on such data are of little use to land managers that make decisions at scales beyond that of traditional population models [@Queenborough2011]. While density-structured approaches may fail when climate covariates are considered, there are other alternatives. For example, @Clark2011 use Forest Inventory and Analysis (FIA) data to parameterize a population model with multiple vital rates and climate dependence. Another example are distributed efforts like PlantPopNet (http://plantago.plantpopnet.com) that will allow researchers to estimate variation around climate responses for widespread species by taking advantage of spatial variation in climate [e.g. @Doak2010]. Lastly, we foresee new approaches on the horizon that leverage photo/video of plots and advanced object recognition algorithms [e.g. @Liu2014] to streamline plant mapping and digitizing efforts.

### Forecasting the future, and the future of forecasting
Our goal was not to make any explicit forecast for the future state of these populations based on predicted climate change. But our results highlight the state of affairs in ecology when it comes to forecasting the impacts of climate change. The analysis we conducted here could be considered, with some exceptions of course, at the forefront of ecological forecasting in terms of the statistical approach employed (hierarchical Bayesian), the type of population model we used (stochastic IPM with parameter uncertainty), and the amount of high quality data we had at our disposal (14 years of individual level data). Yet, model predictions proved so uncertain that any forecast, when bounded with model and parameter uncertainty, would be at best not useful and at worst meaningless. How might we improve on this state of affairs?

First, forecasts could be improved by matching the spatial scale of predictor variables with the spatial scale of observations. One of the major limitations of the models we fit here is that the climate data are at a much larger scale than the individual level observations of plant size. Climate covariates only vary by year, with no spatial variability within years. Thus, even if we fit models to individual level data, we are missing the key interaction point between weather and individual plants [@Clark2011b] because all observations share the same climate covariates. Demographic studies should be designed with at least plot level measurements of climate related variables (e.g., soil moisture).

Second, accurately detecting climate signals will take even longer time series. Recent theoretical work on detecting climate signals in noisy data suggests that even advanced approaches to parameter fitting like LASSO, functional linear models (splines), and Random Forest models require 20-25 year time series (Teller et al., in review). Alternatively, as we suggest above, Teller et al. (in review) also find that matching the scale of the response and predictors improves estimate precision.

Third, ecologists as a community need to get serious about reporting uncertainty. There is a strong culture around explicitly considering model uncertainty, but parameter uncertainty is often ignored. In some cases this is because the easiest statistical methods do no make propogating parameter uncertainty a straighforward task. Even Bayesian approaches that allow integration of model fitting and forecasting [@Hobbs2015] are not simple when using modeling approaches like integral projection models that separate the model fitting and simulation stages [@Rees2009]. However, as we have done here, it is still possible to include parameter uncertainty by drawing parameter values from MCMC iterations, taking care to draw all parameters from the same chain and iteration to account for their correlations. Only by being honest about our forecasts can we begin to produce better ones.

Conclusions
-----------
This work is not a critique of density-structured population models. In some cases and for certain species, population models based on aggregated data may prove useful and unbiased. However, our work here is the first comparison, to our knowledge, of population models based on individual and aggregated data in a climate change context. Our results confirm theoretical arguments [@Clark2011b] and empirical evidence [@Clark2011; @Clark2012] that individual responses are critical to predicting species' responses to climate change. Thus, forecasts from aggregate level models should be viewed with caution and should never be unaccompianed by uncertainty. Given the importance of demographic data and its current difficulty to collect, we advocate for research on modern methods to collect demographic data more efficiently across environmental gradients in space and time.

@@

Acknowledgments
---------------
This work was funded by the National Science Foundation through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040). We thank the original mappers of the permanent quadrats in Montana and the digitizers in the Adler lab, without whom this work would not have been possible. Informal conversations with Stephen Ellner, Giles Hooker, Robin Snyder, and a series of meetings between the Adler and Weecology labs at USU sharpened our thinking. Compute, storage and other resources from the Division of Research Computing in the Office of Research and Graduate Studies at Utah State University are gratefully acknowledged.

### 2015-07-09 manuscript/manuscript.tex
updates ms with new references

where $k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})$ is the population kernal
that describes all possible transitions from size $u$ to $v$ and
$\bar{\bold{w_{j}}}$ is a vector of estimates of average crowding
experienced from all other species by a genet of size $u_j$ and species
$j$. The integral is evaluated over all possible sizes between
predefined lower (\emph{L}) and upper (\emph{U}) size limits that extend
beyond the range of observed genet sizes.

The population kernal is defined as the joint contributions of survival
(\emph{S}), growth (\emph{G}), and recruitment (\emph{R}):

@@

Equilibrium cover from both models was sensitive to climate (Fig. 3a-d).
The IPM projected percent changes in equilibrium cover from -3 - 8\% for
\emph{B. gracilis}, -4 - 3\% for \emph{H. comata}, -15 - 9\% for
\emph{P. smithii}, and -17 - 53\% for \emph{P. secunda}. The QBM
projected opposite and greater percent changes in equilibrium cover for
\emph{B. gracilis} (-63 - 30\%) and \emph{H. comata} (-50 - -18\%; Fig.
3a-b). For \emph{P. smithii}, the QBM projected opposite changes in
equilibrium cover than the IPM, but of similar magnitude (-5 - 6\%; Fig.
3c). \emph{P. secunda} was the only species that the IPM and QBM made
projections of the same sign and somewhat similar magnitude (Fig. 3d).

The response of a population to climate change is a result of the
aggregate effects of climate on individual vital rates. Since the IPM
approach relies on vital rate regressions, we were able to quantify the
sensitivity of each vital rate in isolation and in pairwise
combinations. Species showed similar trends (Fig. 3e-h). Growth was the
most sensitive vital rate for all species, showing a negative response
to increased precipitation, and stronger positive response to increased
temperature, and a mostly positive response when both climate factors
are increased (Fig. 3e-h). \emph{B. gracilis} survival rates were
sensitive to temperature, resulting in an increase in plant cover under
increased temperature (Fig. 3e). In isolation, recruitment and survival
were insensitive to climate factors for \emph{H. comata} (Fig. 3f).
Survival and recruitment of \emph{P. smithii} were both sensitive,
negatively, to temperature and precipitation (Fig. 3g). \emph{P.
secunda} equilibrium cover was sensitive to the climate effects on
survival and recruitment, showing a negative effect on both vital rates
for increased precipition, but a strong positive effect on survival with
increased temperature (Fig. 3h). The climate impact of recruitment on
equilibrium cover was negative for precipitation and temperature
increases (Fig. 3h).

@@

These results lead us to conclude that detailed demographic data is
necessary to forecast climate change impacts on plant populations. This
is unwelcome news since this data is difficult to collect and the models
built on such data are of little use to land managers that make
decisions at scales beyond that of traditional population models
(Queenborough et al. 2011). While density-structured approaches may fail
when climate covariates are considered, there are other alternatives.
For example, ({\textbf{???}}) use Forest Inventory and Analysis (FIA)
data to parameterize a population model with multiple vital rates and
climate dependence. Another example are distributed efforts like
PlantPopNet (\url{http://plantago.plantpopnet.com}) that will allow
researchers to estimate variation around climate responses for
widespread species. Lastly, we foresee new approaches on the horizon
that leverage photo/video of plots and advanced object recognition
algorithms to streamline plant mapping and digitizing efforts.

@@

Our goal was not make any explicit forecast for the future state of
these populations based on predicted climate change. But our results
highlight the state of affairs in ecology when it comes to forecasting
the impacts of climate change. The analysis we conducted here could be
considered, with some exceptions of course, at the forefront of
ecological forecasting in terms of the statistical approach employed
(hierarchical Bayesian), the type of population model we used
(stochastic IPM with parameter uncertainty), and the amount of data we
had at our disposal (14 years of individual-level data). Yet, model
predictions proved so uncertain that any forecast, when bounded with
model and parameter uncertainty, would be at best not useful and at
worst meaningless. How might we improve on this state of affairs?

First, forecasts could be improved by matching the spatial scale of
predictor variables with the spatial scale of observations. One of the
major limitations of the models we fit here is that the climate data are
at a much larger scale than the plot and individual level observations
of plant cover and size. Climate covariates only vary by year, with no
spatial variability within years. Thus, even if we fit models to
individual level data, we are missing the key interaction point between
weather and individual plants (Clark et al. 2011b).

Second, accurately detecting climate signals will take even longer time
series. Recent theoretical work on detecting climate signals in noisy
data suggests that even advanced approaches to parameter fitting like
LASSO, functional linear models (splines), and random forest require
20-25 year time series (Teller et al., in review). Alternatively, as we
suggest above, Teller et al. (in review) also find that matching the
scale of the response and predictors improves estimate precision.

Third, ecologists as a community need to get serious about reporting
uncertainty. There is a strong culture around explicitly considering
model uncertainty, but parameter uncertainty is often ignored. In some
cases this is because the easiest statistical methods do no make
propogating parameter uncertainty easy. Even using Bayesian approaches
that allow straightforward integration of model fitting and forecasting
(Hobbs and Hooten 2015) is not simple when using modeling approaches
like integral projection models that separate the model fitting and
simulation stages (Rees and Ellner 2009). However, as we have done here,
it is still possible to include parameter uncertainty by drawing
parameter values from MCMC iterations, taking care to draw all
parameters from the same chain and iteration to account for their
correlations. Only by being honest about our forecasts can we begin to
produce better ones.

@@

This work is not a critique of density-structured population models. In
some cases and for certain species, population models based on
aggregated data may prove useful and unbiased. However, our work here is
the first comparison, to our knowledge, of population models based on
individual and aggregated data in a climate change context. Our results
confirm theoretical arguments (Clark et al. 2011b) and empirical
evidence (Clark et al. 2011a, 2012) that individual responses are
critical to predicting species' responses to climate change. Thus,
forecasts from aggregate level models should be viewed with caution and
should never be unaccompianed by uncertainty. Given the importance of
demographic data and its current difficulty to collect, we advocate for
research on modern methods to collect demographic data more efficiently
across environmental gradients in space and time. For example, recent
advances in object recognition algorithms (Liu et al. 2014) provide new
opportunities to speed up data collection and digital processing.

@@

This work was funded by the National Science Foundation through a
Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
CAREER award to PBA (DEB-1054040). We thank the original mappers of the
permanent quadrats in Montana and the digitizers in the Adler lab,
without whom this work would not have been possible. Informal
conversations with Stephen Ellner, Giles Hooker, Robin Snyder, and a
series of meetings among the Adler and Weecology labs at USU sharpened
our thinking. Compute, storage and other resources from the Division of
Research Computing in the Office of Research and Graduate Studies at
Utah State University are gratefully acknowledged.

@@

>>>>

where $k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})$ is the population kernel
that describes all possible transitions from size $u$ to $v$ and
$\bar{\bold{w_{j}}}$ is a vector of estimates of average crowding
experienced from all other species by a genet of size $u_j$ and species
$j$. The integral is evaluated over all possible sizes between
predefined lower (\emph{L}) and upper (\emph{U}) size limits that extend
beyond the range of observed genet sizes.

The population kernel is defined as the joint contributions of survival
(\emph{S}), growth (\emph{G}), and recruitment (\emph{R}):

@@

Equilibrium cover from both models was sensitive to climate (Fig. 3a-d).
The IPM projected percent changes in equilibrium cover from -3 to 8\%
for \emph{B. gracilis}, -4 to 3\% for \emph{H. comata}, -15 to 9\% for
\emph{P. smithii}, and -17 to 53\% for \emph{P. secunda}. The QBM
projected opposite and greater percent changes in equilibrium cover for
\emph{B. gracilis} (-63 to 30\%) and \emph{H. comata} (-50 to -18\%;
Fig. 3a-b). For \emph{P. smithii}, the QBM projected opposite changes in
equilibrium cover than the IPM, but of similar magnitude (-5 to 6\%;
Fig. 3c). \emph{P. secunda} was the only species that the IPM and QBM
made projections of the same sign and somewhat similar magnitude (-20 to
14\%; Fig. 3d).

The response of a population to climate change is a result of the
aggregate effects of climate on individual vital rates. Since the IPM
approach relies on vital rate regressions, we were able to quantify the
sensitivity of each vital rate in isolation and in pairwise
combinations. Species showed similar trends (Fig. 3e-h). Growth was the
most sensitive vital rate for all species, showing a negative response
to increased precipitation, and stronger positive response to increased
temperature, and a mostly positive response when both climate factors
are increased (Fig. 3e-h). \emph{B. gracilis} survival rates were
sensitive to temperature, resulting in an increase in plant cover under
increased temperature (Fig. 3e). In isolation, recruitment and survival
were insensitive to climate factors for \emph{H. comata} (Fig. 3f).
Survival and recruitment of \emph{P. smithii} were both sensitive,
negatively, to temperature and precipitation (Fig. 3g). \emph{P.
secunda} equilibrium cover was sensitive to the climate effects on
survival and recruitment, showing a negative effect on both vital rates
for increased precipition, but a strong positive effect on survival with
increased temperature (Fig. 3h). The climate impact of recruitment on
equilibrium cover was negative for precipitation and temperature
increases (Fig. 3h). At least two of three vital rates were sensitive to
climate for each species (Fig. 3).

@@

These results lead us to conclude that detailed demographic data is
necessary to forecast climate change impacts on plant populations. This
is unwelcome news since this data is difficult to collect and the models
built on such data are of little use to land managers that make
decisions at scales beyond that of traditional population models
(Queenborough et al. 2011). While density-structured approaches may fail
when climate covariates are considered, there are other alternatives.
For example, Clark et al. (2011a) use Forest Inventory and Analysis
(FIA) data to parameterize a population model with multiple vital rates
and climate dependence. Another example are distributed efforts like
PlantPopNet (\url{http://plantago.plantpopnet.com}) that will allow
researchers to estimate variation around climate responses for
widespread species by taking advantage of spatial variation in climate
(e.g. Doak and Morris 2010). Lastly, we foresee new approaches on the
horizon that leverage photo/video of plots and advanced object
recognition algorithms (e.g. Liu et al. 2014) to streamline plant
mapping and digitizing efforts.

@@

Our goal was not to make any explicit forecast for the future state of
these populations based on predicted climate change. But our results
highlight the state of affairs in ecology when it comes to forecasting
the impacts of climate change. The analysis we conducted here could be
considered, with some exceptions of course, at the forefront of
ecological forecasting in terms of the statistical approach employed
(hierarchical Bayesian), the type of population model we used
(stochastic IPM with parameter uncertainty), and the amount of high
quality data we had at our disposal (14 years of individual level data).
Yet, model predictions proved so uncertain that any forecast, when
bounded with model and parameter uncertainty, would be at best not
useful and at worst meaningless. How might we improve on this state of
affairs?

First, forecasts could be improved by matching the spatial scale of
predictor variables with the spatial scale of observations. One of the
major limitations of the models we fit here is that the climate data are
at a much larger scale than the individual level observations of plant
size. Climate covariates only vary by year, with no spatial variability
within years. Thus, even if we fit models to individual level data, we
are missing the key interaction point between weather and individual
plants (Clark et al. 2011b) because all observations share the same
climate covariates. Demographic studies should be designed with at least
plot level measurements of climate related variables (e.g., soil
moisture).

Second, accurately detecting climate signals will take even longer time
series. Recent theoretical work on detecting climate signals in noisy
data suggests that even advanced approaches to parameter fitting like
LASSO, functional linear models (splines), and Random Forest models
require 20-25 year time series (Teller et al., in review).
Alternatively, as we suggest above, Teller et al. (in review) also find
that matching the scale of the response and predictors improves estimate
precision.

Third, ecologists as a community need to get serious about reporting
uncertainty. There is a strong culture around explicitly considering
model uncertainty, but parameter uncertainty is often ignored. In some
cases this is because the easiest statistical methods do no make
propogating parameter uncertainty a straighforward task. Even Bayesian
approaches that allow integration of model fitting and forecasting
(Hobbs and Hooten 2015) are not simple when using modeling approaches
like integral projection models that separate the model fitting and
simulation stages (Rees and Ellner 2009). However, as we have done here,
it is still possible to include parameter uncertainty by drawing
parameter values from MCMC iterations, taking care to draw all
parameters from the same chain and iteration to account for their
correlations. Only by being honest about our forecasts can we begin to
produce better ones.

@@

This work is not a critique of density-structured population models. In
some cases and for certain species, population models based on
aggregated data may prove useful and unbiased. However, our work here is
the first comparison, to our knowledge, of population models based on
individual and aggregated data in a climate change context. Our results
confirm theoretical arguments (Clark et al. 2011b) and empirical
evidence (Clark et al. 2011a, 2012) that individual responses are
critical to predicting species' responses to climate change. Thus,
forecasts from aggregate level models should be viewed with caution and
should never be unaccompianed by uncertainty. Given the importance of
demographic data and its current difficulty to collect, we advocate for
research on modern methods to collect demographic data more efficiently
across environmental gradients in space and time.

@@

This work was funded by the National Science Foundation through a
Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
CAREER award to PBA (DEB-1054040). We thank the original mappers of the
permanent quadrats in Montana and the digitizers in the Adler lab,
without whom this work would not have been possible. Informal
conversations with Stephen Ellner, Giles Hooker, Robin Snyder, and a
series of meetings between the Adler and Weecology labs at USU sharpened
our thinking. Compute, storage and other resources from the Division of
Research Computing in the Office of Research and Graduate Studies at
Utah State University are gratefully acknowledged.

@@

Doak, D. F., and W. F. Morris. 2010. Demographic compensation and
tipping points in climate-induced range shifts. Nature 467:959--962.
